---
title: "Bio3415_Lab2_Report: Quality Control And Filtering"
author: "Catherine"
date: "2026-09-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Install fitlong using conda 
Use conda to install the filtlong package a package that allow to filter long reads based on certain criteria such as average Q score, or size.

```{bash, eval=FALSE}
conda install -c bioconda filtlong
```


## Filter The Long Reads
Write a bash code that filters the long read fastq you used earlier with sequali to only keep the longest 75% of the sequences and save the file in the outputs/ directory.

```{bash, eval=FALSE}
filtlong --keep_percent 75 outputs/ERR1539006.fastq > outputs/filtered_reads_ERR1539006.fastq
```


## Run Sequali on the filtered file
```{bash, eval=FALSE}
sequali --outdir outputs/ outputs/filtered_reads_ERR1539006.fastq
```


## Results changed before and after filtration
- After filteration with filtlong, the number of reads and the number of bases decreased because the low-quality read, short fragments, and sequence artifacts has been removed. Consequently, the overall average quality of the remaining read dataset improved significantly, providing a cleaner and more reliable set of data for downstream tasks.

- By this reason, after getting the high-quality after filteration, the result can use for downstream anaylysis such as genome assembly where high accuracy and clean sequence construction are critical to producing reliable results.


## Update github and .gitignore
```{r, eval=FALSE}
*.fastq

*.fastq.gz
```

## Download illumina reads of a Campilobacter jejuni genome sequencing

```{r, eval=FALSE}
#Subset 1
wget -nc -P data/ ftp://ftp.sra.ebi.ac.uk/vol1/fastq/ERR112/040/ERR11203340/ERR11203340_1.fastq.gz

#Subset 2
wget -nc -P data/ ftp://ftp.sra.ebi.ac.uk/vol1/fastq/ERR112/040/ERR11203340/ERR11203340_2.fastq.gz
```



## Plot the sequence quality per base/cycle
```{r, eval=FALSE}
#Load the package
library(Rqc)
#Get the path to the file
folder <- data

#Perform the quality check
#Subset 1
qcRes <- rqc(path = "/Users/cho68/bio3415-lab2/data", pattern = "ERR11203340_1.fastq.gz", openBrowser=FALSE, outdir="outputs")

rqcCycleQualityBoxPlot(qcRes)

#Subset 2
qcRes2 <- rqc(path = "/Users/cho68/bio3415-lab2/data", pattern = "ERR11203340_2.fastq.gz", openBrowser=FALSE, outdir="outputs")

rqcCycleQualityBoxPlot(qcRes2)
```

Analysis: Based on that write at which cycle do you see quality score dropping below 2?

- Based on the graph, at cycle 101, the quality sccore is dropping below 2 



# Plot the sequence content per base per cycle
```{r, eval=FALSE}
#Plot Subset 1
rqcCycleBaseCallsLinePlot(qcRes)

#Plot Subset 2
rqcCycleBaseCallsLinePlot(qcRes2)
```

Analysis: based on that where do you see the highest proportion on Ns?

- Based on the graph, at cycle 5, the highest proportion on Ns occur



# Plot the frequency plot: 
```{r, eval=FALSE}
#Plot Subset 1
rqcReadFrequencyPlot(qcRes)

#Plot Subset 2
rqcReadFrequencyPlot(qcRes2)
```

Analysis: based on that what is the level of redundancy of most seqs (e.g. do most seqs have more than one read)?

- Based on the graph, most of the reads have only one copy. However, there are approximately 10% of reads have up to 3 copies. 



Trim the reads based on the quality scores. trim bases from the 3' end inwards that have Q score <20 (2 marks). 
```{r, eval=FALSE}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
# Biocmanager::install("quasr")

library(QuasR)

# Obtain a list of fastq file paths
fastqFiles <- c("data/ERR11203340_1.fastq.gz", "data/ERR11203340_2.fastq.gz")
                         
outfiles <- paste0("outputs/", basename(fastqFiles))

# Process fastq files
preprocessReads(fastqFiles, outfiles, 
                truncateEndBases=3,
                Lpattern="ACCCGGGA",
                minLength=40)
```


















